On the performance of the MAC model

### 2018-10-29 2018_nips_vigil/introduction.tex
input unit
Mac and S-Mac

\section{Introduction}
Reasoning over visual inputs is a fundamental characteristic of human intelligence.
Reproducing this ability with artificial agents is a challenge that requires learning relations and compositionality (Hu et al. 2017; Johnson et al. 2017b).\\
The Visual Question Answering (VQA) task has been tailored to benchmark those reasoning performances. It is a complex semantic task requiring both natural language processing and visual recognition. It also put to the test the effective use of the short term memory.\\
The recent CLEVR dataset (Johnson et al., 2017a) proposes a challenging multimodal task that requires to answer compositional questions about an image. The dataset is designed with the explicit goal of enabling detailed analysis of visual reasoning.
Solving this problem involves a broad array of skills such as counting, comparison and understanding transitive reasoning.
It also requires perceptual abilities such as recognizing objects, attributes, and spatial relations as well as leveraging commonsense world knowledge (Hudson et al. 2018).
CLEVR is a generated synthetic dataset that allows to introduce variations on a subset of data to test a particular ability such as generalization or transfer learning.\\
To solve this type of reasoning task, many deep learning approaches have been explored although they often struggle to perform well on tasks with a structured and compositional nature (Garnelo et al., 2016; Lake et al., 2017). 
More recent approaches adopt modular structures, resembling the trees of programming languages, that compose neural modules from a fixed predefined collection (Andreas et al., 2016a; Johnson et al., 2017b, Mascharka et al., 2018). However, they rely on externally provided structured representations and functional programs, parsers or expert demonstration, and sometimes require reinforcement learning training schemes (Hudson et al. 2018). 
These models’ structure is inherently rigid and the use of a range of specialized modules weaken their robustness and generalization capacities.\\
Recent work on CLEVR from Hudson et al. 2018;  has introduced  MAC ( Memory, Attention, and Composition). The MAC model has been deliberately designed to decompose the problem into a sequence of attention-based reasoning operations. Its ability to arbitrarily draw a complex reasoning graph while still featuring end-to-end differentiability made its sucess.
MAC doesn’t use specific modules that would make it specifically tailored for the CLEVR dataset. The three general units working in tandem to perform a reasoning step make the robustness of the model.\\
MAC achieves state-of-the-art accuracy on CLEVR and also on the most difficult CLEVR Humans dataset, were questions are written by humans. 
Although the performance of MAC has been proven, It appears difficult to understand what concepts the model is learning. 
Does the model really learn relations between objects? 
How does the model represent those relations and reasoning steps? 
The interpretability of the MAC model seems yet to be fully defined.
Is the model representing notions and concepts like objects attributes?\\
In this work, we first question the complexity of the MAC model and explore simplification possibilities. We propose a new set of equations that aims to simplify the model and improve training performance. 
We also put to proof the MAC model on the CLEVR CoGenT dataset. The CoGenT dataset has been designed to highlight transfer learning abilities and give a better sense of interpretability. By mixing attributes over two sub-datasets, the CLEVR CoGenT task allow us to see what relations and concepts the model has really learned. We show how MAC performs on CLEVR CoGenT and compare it to state-of-the-art models.
and propose a study about the model failures. This work points out some of the MAC model weaknesses, in particular its failure to differentiate concepts of colors and shape. We finally discuss some improvement possibilities.

>>>>

\section{Introduction}
Reasoning over visual inputs is a fundamental characteristic of human intelligence.
Reproducing this ability with artificial agents is a challenge that requires learning relations and compositionality (Hu et al. 2017; Johnson et al. 2017b).

The Visual Question Answering (VQA) task has been tailored to benchmark those reasoning performances. It is a complex semantic task requiring both natural language processing and visual recognition. It also put to the test the effective use of the short term memory.

The recent CLEVR dataset (Johnson et al., 2017a) proposes a challenging multimodal task that requires to answer compositional questions about an image. The dataset is designed with the explicit goal of enabling detailed analysis of visual reasoning.
Solving this problem involves a broad array of skills such as counting, comparison and understanding transitive reasoning.
It also requires perceptual abilities such as recognizing objects, attributes, and spatial relations as well as leveraging commonsense world knowledge (Hudson et al. 2018).
CLEVR is a generated synthetic dataset that allows to introduce variations on a subset of data to test a particular ability such as generalization or transfer learning.

To solve this type of reasoning task, many deep learning approaches have been explored although they often struggle to perform well on tasks with a structured and compositional nature (Garnelo et al., 2016; Lake et al., 2017). 
More recent approaches adopt modular structures, resembling the trees of programming languages, that compose neural modules from a fixed predefined collection (Andreas et al., 2016a; Johnson et al., 2017b, Mascharka et al., 2018). However, they rely on externally provided structured representations and functional programs, parsers or expert demonstration, and sometimes require reinforcement learning training schemes (Hudson et al. 2018). 
These models’ structure is inherently rigid and the use of a range of specialized modules weaken their robustness and generalization capacities.

Recent work on CLEVR from Hudson et al. 2018;  has introduced  MAC ( Memory, Attention, and Composition). The MAC model has been deliberately designed to decompose the problem into a sequence of attention-based reasoning operations. Its ability to arbitrarily draw a complex reasoning graph while still featuring end-to-end differentiability made its sucess.
MAC doesn’t use specific modules that would make it specifically tailored for the CLEVR dataset. The three general units working in tandem to perform a reasoning step make the robustness of the model.

MAC achieves state-of-the-art accuracy on CLEVR and also on the most difficult CLEVR Humans dataset, were questions are written by humans. 
Although the performance of MAC has been proven, It appears difficult to understand what concepts the model is learning. 
Does the model really learn relations between objects? 
How does the model represent those relations and reasoning steps? 
The interpretability of the MAC model seems yet to be fully defined.
Is the model representing notions and concepts like objects attributes?

In this work, we first question the complexity of the MAC model and explore simplification possibilities. We propose a new set of equations that aims to simplify the model and improve training performance. 
We also put to proof the MAC model on the CLEVR CoGenT dataset. The CoGenT dataset has been designed to highlight transfer learning abilities and give a better sense of interpretability. By mixing attributes over two sub-datasets, the CLEVR CoGenT task allow us to see what relations and concepts the model has really learned. We show how MAC performs on CLEVR CoGenT and compare it to state-of-the-art models.
and propose a study about the model failures. This work points out some of the MAC model weaknesses, in particular its failure to differentiate concepts of colors and shape. We finally discuss some improvement possibilities.

### 2018-10-29 2018_nips_vigil/mac.tex
output unit
Mac and S-Mac

\section{Description of MAC model}

A MAC network is an end-to-end differentiable architecture primed to perform an explicit multi-step reasoning process, by stringing together p recurrent MAC cells, each responsible for performing one reasoning step. Given a knowledge base K (for VQA, an image) and a task description q (for VQA, a question), the model infers a decomposition into a series of p reasoning operations that interact with the knowledge base, iteratively aggregating and manipulating information to perform the task at hand. It consists of three components: (1) an input unit, (2) the core recurrent network, composed out of p MAC cells, and (3) an output unit, all described below.

\begin{figure}[htbp]
	\centering
	\includegraphics[width=0.7\textwidth]{img/mac_cell.pdf}
	\caption{The Mac Cell. The MAC recurrent cell consists of a control unit, read unit, and write unit, that operate over dual control and memory hidden states. The control unit successively attends to different parts of the task description (question), updating the control state to represent at each timestep the reasoning operation the cell intends to perform. The read unit extracts information out of a knowledge base (here, image), guided by the control state. The write unit integrates the retrieved information into the memory state, yielding the new intermediate result that follows from applying the current reasoning operation.}
	\label{fig:core_concepts}
\end{figure}

>>>>

\section{MAC network and our proposed simplification}

The MAC network~\cite{hudsonManning18} is a recurrent model that performs reasoning 
using multiple steps, where each step involves sequentially analyzing a part of the question following which attention on the image is changed suitably.
At the end, a suitable answer is given based on this composite reasoning.
To improve the accuracy of prediction, both the questions and images
are preprocessed suitably. For the question,  a word-embedding is first used to
transformed the question into a sequence of vectors which is then passed through
a bidirectional LSTM to produce a sequence of hidden state pairs. Each such pair 
corresponds to a unique word in the question and is called the 
\emph{contextual embedding} for that word.
Similarly each image is initially processed using a pre-trained network (using ResNet) followed by two trainable CNN layers.

\begin{figure}[htbp]
	\centering
	\includegraphics[width=0.7\textwidth]{img/mac_cell.pdf}
	\caption{The Mac Cell. The MAC recurrent cell consists of a control unit, read unit, and write unit, that operate over dual control and memory hidden states. }
	\label{fig:mac_cell}
\end{figure}

### 2018-10-29 2018_nips_vigil/smac.tex
Mac and S-Mac

\section{Simplified MAC model}

Idealized equations:
\subsection*{Control Unit}

>>>>

\subsection{Simplified MAC network}
Our proposed modification to the MAC network is based on two heuristic
simplifications of the equations governing a MAC cell. 
First, we observed that when looking at the equations of different units
together as a whole, we notice that there are several affine transformations
that are applied in sequence (with no activation in-between).
Therefore they can actually be combined to form a single affine transformation.
The second observation uses the following heuristic argument:
suppose there is a single affine transformation which preserves the number
of dimensions. Then one can assume that the transformation is invertible
so as to avoid a loss in the information. Applying this principle to 
equations in the read and write unit results in eliminating an affine
operation of the knowledge base in the recurrent network. Instead we compute
this operation on the knowledge base right after we apply the 2 
trainable CNN layers (i.e. prior to the reasoning steps). This results
in a significant savings in training time, as our experiments demonstrate
below, without incurring too much drop in accuracy.

\subsection*{Control Unit}
